\chapter{Survey of current SoP positioning systems}
\label{s_sop}
This chapter contains a survey of current research on the topic of SoP positioning. Firstly, two positioning systems based on the Doppler method are presented for comparison. Then, a survey of state-of-the art experimental SoP positioning systems is conducted, and for each research effort, the methods and results are summarised. Finally, some general conclusions are drawn and the results achieved by the surveyed systems are compared to dedicated positioning systems - the Transit and GNSS systems.



\section{Doppler method positioning system}
\label{s_sop_dopp_systems}
This section briefly discusses two dedicated positioning systems, the Transit and Argos, which use the Doppler method. It is meant to provide a basis of comparison when evaluating the SoP systems which is more appropriate than comparing against GNSS.


\subsection{Transit}
The Navy Navigation Satellite System, commonly referred to as Transit or NAVSAT, was the first operational satellite navigation system. The system was used to provide accurate position updates for the INS equipment aboard Polaris nuclear ballistic missile submarines. Later, civilian use cases such as fishing boat navigation emerged. The system was operational from 1964 to 1996. The system used the Doppler method for positioning, the details of which are described in \autoref{s_pos_curve_fit_method}. The transmission frequencies were \qtylist{150;400}{MHz} (two to allow for compensation of ionospheric refraction) and the system also provided a data message used for clock synchronisation. The performance of the system exceeded that of all of the early large radio systems - the RMSE of a position fix of a static naval user was \qtyrange{40}{100}{m} and the time to fix was \qtyrange{10}{16}{min}, but could be as low as \qty{2}{min} with lower accuracy\cite{sat16}.
% Neměl byste, prosím, ten zdroj 14] někde stažený/online? Celkem rád bych jej získal :-) 
%% je to "sat16" na OwnCloudu

\subsection{Argos}
Argos is a global satellite-based location and data collection system dedicated to studying and protecting Earth environment, operated through a multinational research agency collaboration. It allows any mobile object, usually a tracked animal, equipped with a compatible transmitter terminal to be located across the world\cite{sat17}. Unlike all other systems discussed, the actual position determination is done at the level of the satellite, which receives a signal (which may also contain sensor data) from the terminal and calculates its position.

Argos terminal messages are transmitted at a frequency of \qty{401650(30)}{kHz}. Calculation of terminal position is done through the Doppler shift method, using a classical nonlinear least squares estimation or a multiple-model Kalman filter. The positioning accuracy typically ranges from the order of \qty{e1}{km} to less than \qty{250}{m}, depending on the number of terminal messages received during a satellite pass\cite{sop09}. Algorithms based on animal movement models are used to resolve the symmetrical problem of Doppler positioning\cite{sat17}.



\section{Experimental SoP positioning systems}
This section contains a survey of select experimental SoP positioning systems, each with a brief description of the implementation and experimental results. All of these system use the Doppler-shift method.

In \cite{sop07}, it is demonstrated using analysis and simulation that it is possible to obtain quick\footnote{Definition of "quick" is not provided, but it is stated that it is significantly less than \qty{16}{\s}} positioning using Doppler frequency measurements of signals from one or two satellites of the Globalstar constellation (forward down-link carrier frequency of \qty{2500}{\mega\hertz}). The error in the horizontal plane of such positioning is expected to be less than \qty{9}{\km} in \qty{90}{\percent} of cases using one satellite. Using two satellites, the error is expected to be \qty{1.4}{\km} in \qty{90}{\percent} of cases. This is without any advanced signal processing or data fusion.

In \cite{sop01}, a method of positioning is devised for use as a secondary aircraft navigation system in GPS-denied areas using Iridium NEXT satellites and an SDR. Circular Error Probable (CEP) of \qtyrange{0.2}{2}{\km} in dynamic mode is demonstrated without using significant computational resources. The system is based on receiving STL, Ring Alert and MSG Iridium NEXT signals on \qty{1626.104}{\mega\hertz}, \qty{1626.27}{\mega\hertz} and \qty{1626.44}{\mega\hertz}, respectively, and then fusing the Doppler shift data with IMU data using Kalman filtering.

In \cite{sop04}, a Starlink-based system is designed, based on Doppler measurements of a tone band centred on \qty{11 325}{\mega\hertz}. The system tracked the tone signal of up to six satellites by FLL-assisted-PLL. Extended Kalman filtering was used. The system accuracy was \qty{375}{\metre}. The team determined that the most significant error sources were the receiver (user) clock drift, the user time error and the satellite position error. Notably, the satellite clock drift was not determined to be significant.

In \cite{sop05}, a system based on the fusion (using EKF) of Iridium Doppler shift measurements, INS and magnetometer data is proposed and simulated for a highly dynamic application. The simulated Iridium signal was centred on \qty{1626.25}{\mega\hertz}. The arrival position error\footnote{That is, the error in position after travelling along a predefined trajectory in the simulation} in \qty{95}{\percent} of cases was approximately \qty{300}{\metre} for the best case, and \qty{4}{\km} for the intermediate case. The main difference between the two cases was the accuracy of orbital data.

In \cite{sop08}, a framework is presented which estimates the LEO satellite states along with the states of a navigating vehicle using Orbcomm satellite signals. The system fuses INS with Doppler measurements of a signal in VHF downlink band of \qtyrange{137}{138}{\mega\hertz} using EKF. The system was tested using an UAV, which had GPS signal available for \qty{90}{\s} and then turned off for \qty{30}{\s}. The final position error was \qty{8.8}{\m}, compared with INS-only error of \qty{31.7}{\m}. Static positioning without the use of GNSS was not performed.

In \cite{sop10}, a framework employing EKF to estimate user position from Doppler measurement of multiple generic LEO satellites was proposed. Simulation showed a RMSE in position from approx. \qty{170}{\m} for 5 satellites over one minute to \qty{11.5}{\m} for 25 satellites over four minutes. Experimental run with 2 Orbcomm satellites (broadcasting at \qty{137.3125}{\mega\hertz} and \qty{137.25}{\mega\hertz} respectively) over one minute showed a RMSE of \qty{360}{\m}.

In \cite{sop11}, and Iridium-based system is proposed and tested. The system uses the Ring Alert and Primer message signals to capture Doppler shift data, which are then processed using Kalman filtering and the least squares method. The system accuracy can be as much as \qty{22}{\m} after \qty{24}{\hour} of measurement under an open sky. For \qty{30}{\min} under open sky, the accuracy was shown to be \qty{46}{\m}. Furthermore, the system accuracy in GPS-hostile conditions (in a dense forest) was shown to be\footnote{The experiment time is not stated, but can be inferred to be about \qty{30}{\min}} \qty{108}{\m}, whereas GPS did not provide a solution.

In \cite{sop03}, two LEO constellations (Iridium NEXT and Orbcomm) are used to determine position. Three-dimensional RMSE achieved over \qty{30}{\s} using a single constellation (Orbcomm) is approx. \qty{0.76}{\km}, whereas using both constellations the RMSE is \qtyrange{0.22}{0.18}{\km} depending on the received signal type. Accuracy when measuring messaging bursts was better than for Ring Alert bursts. EKF was utilised in the signal processing. In the receiver architecture, each constellation has a dedicated acquisition chain, including separate antennas, amplifiers, filters and signal processing.


\subsection{Summary}
The research outlined above is summarised in \autoref{t_sop_survey_summary}. Most research focused on using Iridium NEXT satellites as signal source, followed by Orbcomm. The accuracy of position determination was mostly in the order of \qtyrange{e2}{e3}{\m} and improved significantly both with measurement time and with the number of measured satellites. In one case the performance was shown to exceed the one of existing GNSS systems. Furthermore, the surveyed research demonstrated the possibility of using SoP positioning both in static and dynamic applications. Notably, all of the surveyed experiments used Kalman filtering for fusion of multiple navigation system data or error estimation, albeit the actual implementation varied greatly.

\begin{table}
\caption[Summary of selected existing SoP positioning systems]{Summary of selected existing SoP positioning systems (* denotes a simulated result)}
\label{t_sop_survey_summary}
\hspace*{-2cm}
\centering
\begin{tabular}{p{0.2\linewidth}p{0.15\linewidth}lllll}
Constellation          & Signal                             & Mode       & Accuracy (2D)                  & Exp. time               & Source  \\ \hline
Globalstar             & \qty{2500}{\mega\hertz}            & static     & \qtyrange{1.4}{9}{\km}*        & few s                   & \cite{sop07}     \\
Iridium NEXT           & STL, Ring Alert, MSG               & dynamic    & \qtyrange{0.2}{2}{\km}         & not found               & \cite{sop01}     \\
Starlink               & \qty{11 325}{\mega\hertz}          & static     & \qty{375}{\metre}              & \qty{330}{\s}           & \cite{sop04}     \\
Generic, Orbcomm       & \qty{137}{\mega\hertz}             & static     & \num{11.5}* to \qty{360}{\m}   & \qtyrange{1}{4}{\min}   & \cite{sop10}     \\
Iridium NEXT           & Ring Alert, Primer                 & static     & \qtyrange{46}{108}{\m}         & \qty{30}{\min}          & \cite{sop11}     \\
Iridium NEXT, Orbcomm  & MSG, Ring Alert                    & static     & \qtyrange{0.18}{0.76}{\km}     & \qty{30}{\s}            & \cite{sop03}     \\
\multicolumn{6}{l}{\textit{Note: Results from \cite{sop05} and \cite{sop08} are not in this table, as the research concerns}} \\
\multicolumn{6}{l}{\textit{SoP positioning coupled with INS in a dynamic application}} 
\end{tabular}
\end{table}



\section{Comparison to dedicated positioning systems}
\label{s_sop_comparison}
To compare the SoP positioning systems, a representative system needs to be created first. The considered parameters are the system positioning accuracy and the measurement time required to achieve a position fix (TTF, time to fix).

There are too few systems to use standard statistical methods, and the parameters of the systems are too spread apart to be represented by just one system. Furthermore, the systems with better accuracy generally take much longer to acquire a fix. Therefore, two systems were considered for comparison - "SoP A", which represents a more accurate albeit slower system, and "SoP B", which represents the opposite. The SoP A has an accuracy of \qty{200}{m} and time to fix of \qty{600}{s}, whereas the SoP B parameters are \qty{700}{m} and \qty{120}{s}, respectively.

The parameters of the different GNSS systems are quite similar at the level of SoP systems\cite{sop23}, therefore GPS will be used as a representative example. No navigation aiding, such as SBAS, will be considered, as this approximately matches the operation of SoP systems. However, even within GPS the system parameters differ greatly. Geodetic GPS receivers are accurate to a few centimetres, but require up to \qty{30}{min} to obtain a fix with such accuracy\cite{sop23}. On the other hand, commercial-grade receivers are accurate on the level of metres and require no more than few minutes to fix from a cold start\footnote{GPS cold start matches the operation of SoP systems}\citep{pos04}{33}. To represent GPS, a system with an accuracy of \qty{3}{m} and time to fix of \qty{120}{s} will be considered.

Additionally, the Transit system (\qty{70}{m}, \qty{13}{min}) will be considered (see \autoref{s_sop_dopp_systems}). The Argos system will not be considered, as its parameters vary greatly and its operation is different.

The systems are summarised in \autoref{t_sop_comparison}. Clearly, GPS performs best in terms of accuracy. Interestingly though, the time of fix of the faster SoP system (SoP B) is similar to that of cold start GPS. Compared to Transit, the SoP systems are faster, and the difference in accuracy is much less, albeit still not in favour of the SoP systems. 

The findings of this comparison are hardly surprising, as even the Doppler method requires time synchronisation, which is not provided by SoP signals, and neither are precise ephemerides to increase the accuracy of satellite orbit determination or any other navigation aids. Furthermore, the SoP systems are based on systems which were never intended for precision navigation, such as the TLEs, or indeed the communication satellites themselves - the clock accuracy and therefore the transmission frequency is not assured on a navigation level.

On the other hand, some research showed that the SoP system is capable of acquiring a fix in conditions when GPS is not; and the demonstrated parameters are sufficient for some applications, e.g. emergency location (see the discussion in \autoref{s_int_use_cases} for more details).


\begin{table}
    \centering
    \begin{tabular}{l|ll|ll}
        System  & Accuracy (m) & $\times$ best & Time to Fix (s) &  $\times$ best \\ \hline
        GPS     & \textbf{3}   &             1 &    \textbf{120} & 1   \\
        Transit & 70           &            23 &             780 & 6.5 \\
        SoP A   & 200          &            67 &             600 & 5   \\
        SoP B   & 700          &           233 &    \textbf{120} & 1   \\
    \end{tabular}
    \caption{Comparison of SoP and dedicated positioning systems}
    \label{t_sop_comparison}
\end{table}

% Tak s tou volbou metriky na srovnání jste se vyrovnal skvěle!
% Jen ta informace o konvergenci RTK měření u GPS  - 30 minut. To už je trochu starší údaj. Dnes je to i do minuty. Ale klidně tam dejte "up to", nebo že historicky to je i až 30 minut.
%% opravil jsem
%% Díky!